Technique for \earlywarning\ filtering for compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-29 introduction.tex
Removed more commented out text
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1184 ccdd05d5-56c1-461a-bf98-85f6876523b4
Removed citation to Numerical Recipes, since it is cited again later in the text (and overlap-save is a standard algorithm)
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1185 ccdd05d5-56c1-461a-bf98-85f6876523b4

In October 2010, \LIGO{}\footnote{\url{http://www.ligo.org/}} completed its
sixth science run (S6) and Virgo\footnote{\url{http://www.ego-gw.it/}}
completed its third science run (VSR3).  While both \LIGO{} detectors and Virgo
were operating, several all-sky detection pipelines operated in a low-latency
configuration to send astronomical alerts, namely \textsc{mbta}, Coherent
WaveBurst, and Omega \citep{HugheyGWPAW2011, S6lowlatency}. The S6 analyses
achieved latencies of 30--60 minutes, which were dominated by a human vetting
process. Candidates were sent for \EM{} followup to several telescopes; Swift,
\textsc{rotse}, \textsc{tarot}, and Zadko \citep{kanner2008, HugheyGWPAW2011}
imaged some of the most likely sky locations.  \textsc{mbta} achieved the best
\GW{} trigger-generation latencies of 2--5 minutes.  
%We assume that in the
%advanced detector era the vetting process will be automated, so current \GW{}
%search methodology and telescope actuation would dominate latency.

@@

This work will focus on reducing the latency of trigger production.  Realizing
advance detection of \CBC{}s will require striking a balance between latency
and throughput of filtering \GW\ data. \CBC{} searches consist of banks of
matched filters, or cross-correlations between the data stream and a bank of
nominal ``template'' signals.  There are many different implementations of
matched filters, but most have high throughput at the cost of high latency, or
low latency at the cost of low throughput.  The former are epitomized by the
overlap-save algorithm % \citep{numerical-recipes-chapter-13}
for frequency domain (\FD) convolution, currently the preferred method in \GW{}
searches.  The most obvious example of the latter is the time domain (\TD)
convolution, which is latency-free.  However, its computational complexity is
quadratic in the length of the templates, so it is prohibitively expensive for
long templates.

>>>>

In October 2010, \LIGO{}\footnote{\url{http://www.ligo.org/}} completed its
sixth science run (S6) and Virgo\footnote{\url{http://www.ego-gw.it/}}
completed its third science run (VSR3).  While both \LIGO{} detectors and Virgo
were operating, several all-sky detection pipelines operated in a low-latency
configuration to send astronomical alerts, namely \textsc{mbta}, Coherent
WaveBurst, and Omega \citep{HugheyGWPAW2011, S6lowlatency}. The S6 analyses
achieved latencies of 30--60 minutes, which were dominated by a human vetting
process. Candidates were sent for \EM{} followup to several telescopes; Swift,
\textsc{rotse}, \textsc{tarot}, and Zadko \citep{kanner2008, HugheyGWPAW2011}
imaged some of the most likely sky locations.  \textsc{mbta} achieved the best
\GW{} trigger-generation latencies of 2--5 minutes.

@@

This work will focus on reducing the latency of trigger production.  Realizing
advance detection of \CBC{}s will require striking a balance between latency
and throughput of filtering \GW\ data. \CBC{} searches consist of banks of
matched filters, or cross-correlations between the data stream and a bank of
nominal ``template'' signals.  There are many different implementations of
matched filters, but most have high throughput at the cost of high latency, or
low latency at the cost of low throughput.  The former are epitomized by the
overlap-save algorithm
for frequency domain (\FD) convolution, currently the preferred method in \GW{}
searches.  The most obvious example of the latter is the time domain (\TD)
convolution, which is latency-free.  However, its computational complexity is
quadratic in the length of the templates, so it is prohibitively expensive for
long templates.
